6.867 Problem Set 1

### 2015-12-10 Project/Latex/Paper.tex
[project] final paper r evisions.

\section{Data}
\subsection{NYT Internal Web Traffic Data}
Our NYT internal web traffic dataset is a record of all individual user activity on the NYT website covering the period of April 3rd, 2013 to October 31st, 2013. Each time a user\footnote{in this case, a ``user" is uniquely defined by a device/browser id. So while the same person might have multiple devices or may use multiple browsers, the NYT backend treats each device/browser combination as a unique ``user" even though in reality its all the same person. In some cases, we are able to link various id's together if the person happens to register an official user account on the NYT website and then logs into her account from multiple devices/browsers.} moves from one page to another on the NYT website, this activity is captured as an individual JSON object. This data is incredibly detailed and can potentially tell us who accessed what page when and from what location. In addition, it tracks the amount of time a user stays on a webpage and path a user took to get to her current webpage. Overall, this dataset is over 20 terabytes in size and contains over 3 billion page views. It's important to note that not all page views are page views of content we care about. For example, searches or changes in user account settings are also recorded in our dataset. Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013, with plans of extending our analysis to our entire dataset in the future. We further restrict our dataset to only include articles or blogposts, since these are the pieces of content that contain readily extractable text data. For the sake of brevity, we use the term ``articles" to refer to both articles and blogposts unless we explicitly state otherwise.

We conducted a first pass through the data to simply to obtain a list of URLs. After cleaning up the url data to ensure each url mapped uniquely to a particular piece of content, we were left with a total of 6682 URLs. We then parses all the web data for the month of August and the first week of September, counting the number of impressions each URL received. In order to make an apples to apples comparison between articles, we only count the number of page views received in the 7 days immediately following publication, since an article that's been out longer should have more page views in expectation. Given the tendency for viewership of an article to drop off sharply soon after publication (as recency is an important factor in news readership), our 7-day measure generally represents the vast majority (well over 90\%) of total page views that an article receives\footnote{at least for a reasonable stretch of time}. Even after all this subsampling, our data still consists of 248,161,455 page views\footnote{Though one video by PSY completely crushes this number}. As seen in Figure 1 below, the distribution of page views is highly skewed with very heavy tails. After applying a log transformation (as seen in Figure 2), our distribution looks considerably more normal.

@@

\subsection{Article Perplexity}
In order to determine the perplexity score, we first need to build some language model that gives us the probabilities of each word. While perplexity typically is a measure of how well a probability distribution can predict a sample, in our context, we interpret perplexity essentially as a measure of article “uniqueness”. The argument here is that if our language model can’t predict the language used in article very well, then the language used in the article is atypical relative to the corpus used to build the language model. Hence, given some language model, an article's perplexity is given by:
\begin{equation}
	2^{-n\cdot\sum_{i=1}^n \log p(w_i)}
\end{equation}
where $n$ is the length of the article, and $p(w_i)$ is the probability of the $i$-th word in the article. We think that perplexity might have some predictive power since people generally have a preference for novelty. If many news articles about the same story are all using highly similar language, an article that covers the story using atypical language is likely unique in some way or another which may drive people to read it more or less. For this section we generally follow the 6.864 Lecture Notes \cite{6864Lec23}.

As for out paper, we construct a couple of different language models. First, we build a simple bi-gram language to use as a baseline. We also build more sophisticated word vector based n-gram neural network language models.

We split our articles into training (~70\%), validation (~15\%), and test corpora (~15\%). In order to keep the size of our vocabulary relatively manageable, we ignore any case sensitivities (so ``Cat" is the same as ``cat"). Furthermore, we only include a word if it appears at least 5 times. Words that don't make this cutoff are mapped to a generic ``rare word" indicator. Lastly, we also map any numbers (that is numbers comprised of digits, not numbers written with words) to a generic ``number" indicator. Ultimately, this leaves us with a vocabulary size $|V|$ of 29359. To estimate a bigram model, we simply need to compute the counts in our training corpus. Specifically, the probability of some word $w_i$ conditional on its preceding word $w_{i-1}$ is given by:
\begin{equation}
p(w_i|w_{i-1}) = \frac{\textnormal{count}(w_{i-1},w_i)}{\textnormal{count}}
\end{equation}
However, its reasonable to expect that there might be bigrams in the development or test corpora that aren't observed in the training corpus considering that for our vocabulary size, there are nearly 900 million unique bigrams and our training data contains just over 3 million observations. If this is the case, then any article with a bigram unobserved in the training corpus would be assigned a predicted likelihood of 0. Needless to say, this is very bad. In order to avoid this issue, we apply a technique called add-$\alpha$ smoothing. Add-$\alpha$ smoothing just adds $\alpha$ to each cell in the probability table. Hence, after smoothing, no bigram, given a fixed vocabulary $V$, will ever have 0 probability. This changes our our estimated word probabilities to:
\begin{equation}
p(w_i|w_{i-1}) = \frac{\textnormal{count}(w_{i-1},w_i) + \alpha}{\textnormal{count} + \alpha |V|}
\end{equation}
Essentially, this smoothing is acting like a probablistic Robin Hood, it takes probability mass from the ``rich" and distributes it to the ``poor". Additionally, this smoothing actually has a nice Bayesian interpretation as well. For the bigram language model case, we can say that words follow a multinomial distribution conditional on its preceding word. The smoothed language model is equivalent to the posterior of horizontally stacked multinomial likelihoods with a symmetric Dirichlet prior with a parameter value of $\alpha$.

@@

Here we see exactly what we should expect. The NALL on the training corpus is lower than the NALL on the validation corpus. Furthermore, the NALL on our training corpus monotonically increases as we increase $alpha$ while the NALL on the validation corpus initially decreases hits a minimum, and starts to increase. For our language model, the optimal $\alpha$ is 0.004. This level of smoothingproduces the following NALLs for our 3 corpora:

@@

One criticism of standard n-gram language models is that they are rather sensitive to the training data. Suppose that two words generally have fairly interchangeable use cases (that is the sets of words that tend to precede these words tend to be highly similar). For example, consider ``coffee" and ``tea". Further suppose that for whatever reason, the phrase ``drink coffee" is predominantly featured in the training data but ``drink tea" is not. Then a standard bigram language model would assign a very low probability to ``tea" if it were preceded by the word ``drink" even if the in the training data contains many instances in which they are fairly interchangeable (``buy coffee/tea", ``brew coffee/tea", etc.).

Hence, we build word vector n-gram neural network models to see if we can achieve better performance. By using dense word vectors rather than one-hot encodings to represent words, we're able capture the underlying similarity of words and their meanings. For example, ``coffee" is more similar to ``tea" that it is to ``car". Rather astonishingly, these word vector values can be trained through back propagation just like the weights in a neural network just as long as they're randomly initiated!

 In particular, we train both a bigram neural network language model and a trigram neural network language model. In both cases, we use word vectors of length 10 meaning that the input layer of the bigram NN is a 10 (a single word) and the input layer of the trigram is 20 (2 words concatenated together). This input layer then maps to the single hidden layer with 10 hidden units through a tanh activation function. Finally, the hidden layer is then connected to the output layer which is has size $|V|$ through a softmax activation function. We use stochastic adaptive gradient descent so that the errors can be properly backporgated to the word vectors. We purposefully keep our network relatively small since our data set is rather large and training even this small neural network takes considerable time.

@@

 We adapted our code so it saved a useable snapshot of the network after each full pass through the data. Unfortunately, as a result of an oversight in the code, we forgot to store the NALL of training data in this snapshot. All is not lost since we remembered to at least store the NALL on the validation corpus. We are highly confident our neural network is working as intended since we see we that the NALL decreases every pass on the validation corpus in Figure \ref{fig:perp_hist} below.

@@

It's worth taking the time to discuss Figure \ref{fig:mse_plot}, which shows the training and holdout MSE for various values of $\lambda$, in slightly more depth. There are a few things in this plot worth discussing. First, note that the validation MSE is consistently higher than the training data MSE, which is consistently higher than the test data MSE. Given the (relatively) small size of our dataset (6,687 observations), this is likely due to the sampling we used to separate our data into training, validation, and test data. However, we don't expect this effect the validity of our cross validation.

@@

Tshere is ample room to improve our sentiment analysis methodology. The training dataset generated using mechanical turk is relatively small (200 training articles). Because of this, there may exist many words that have a strong probability of appearing in an article conditional on sentiment that simply do not appear in our training set. With more time (and money!), we could label more data and improve the accuracy of our model. Furthermore, the model currently skips words that do not occur in the training corpus. An extension of our Naive Bayes implementation could use a method such as Laplace smoothing, so as to not simply ignore words we haven't seen in our training data.

>>>>

\section{Data}
\subsection{NYT Internal Web Traffic Data}
Our NYT internal web traffic dataset is a record of all individual user activity on the NYT website covering the period of April 3rd, 2013 to October 31st, 2013. Each time a user\footnote{In this case, a ``user" is uniquely defined by a device/browser id. So while the same person might have multiple devices or may use multiple browsers, the NYT backend treats each device/browser combination as a unique ``user" even though in reality its all the same person. In some cases, we are able to link various id's together if the person happens to register an official user account on the NYT website and then logs into her account from multiple devices/browsers.} moves from one page to another on the NYT website, this activity is captured as an individual JSON object. This data is incredibly detailed and can potentially tell us who accessed what page when and from what location. In addition, it tracks the amount of time a user stays on a webpage and path a user took to get to her current webpage. Overall, this dataset is over 20 terabytes in size and contains over 3 billion page views. It's important to note that not all page views are page views of content we care about. For example, searches or changes in user account settings are also recorded in our dataset. Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013, with plans of extending our analysis to our entire dataset in the future. We further restrict our dataset to only include articles or blogposts, since these are the pieces of content that contain readily extractable text data. For the sake of brevity, we use the term ``articles" to refer to both articles and blogposts unless we explicitly state otherwise.

We conducted a first pass through the data to simply to obtain a list of URLs. After cleaning up the url data to ensure each url mapped uniquely to a particular piece of content, we were left with a total of 6,682\footnote{If we had just included a few more URLs, we could have had 6,867 observations!} URLs. We then parses all the web data for the month of August and the first week of September, counting the number of impressions each URL received. In order to make an apples to apples comparison between articles, we only count the number of page views received in the 7 days immediately following publication, since an article that's been out longer should have more page views in expectation. Given the tendency for viewership of an article to drop off sharply soon after publication (as recency is an important factor in news readership), our 7-day measure generally represents the vast majority (well over 90\%) of total page views that an article receives\footnote{at least for a reasonable stretch of time}. Even after all this subsampling, our data still consists of 248,161,455 page views\footnote{Though one video by PSY completely crushes this number}. As seen in Figure 1 below, the distribution of page views is highly skewed with very heavy tails. After applying a log transformation (as seen in Figure 2), our distribution looks considerably more normal.

@@

\subsection{Article Perplexity}
In order to determine the perplexity score, we first need to build some language model that gives us the probabilities of each word. While perplexity typically is a measure of how well a probability distribution can predict a sample, in our context, we interpret perplexity essentially as a measure of article “uniqueness”. The argument here is that if our language model can’t predict the language used in article very well, then the language used in the article is atypical relative to the corpus used to build the language model. Hence, given some language model, an article's perplexity is given by:
\begin{equation}
	2^{-n\cdot\sum_{i=1}^n \log p(w_i)}
\end{equation}
where $n$ is the length of the article, and $p(w_i)$ is the probability of the $i$-th word in the article. We think that perplexity might have some predictive power since people generally have a preference for novelty. If many news articles about the same story are all using highly similar language, an article that covers the story using atypical language is likely unique in some way or another which may drive people to read it more or less. For this section we generally follow the 6.864 Lecture Notes 2 and 3 \cite{6864Lec23}.

As for our paper, we construct a couple of different language models. First, we build a simple bigram language to use as a baseline. We also build more sophisticated word vector based n-gram neural network language models.

We split our articles into training (~70\%), validation (~15\%), and test corpora (~15\%). In order to keep the size of our vocabulary relatively manageable, we ignore any case sensitivities (so ``Cat" is the same as ``cat"). Furthermore, we only include a word if it appears at least 5 times. Words that don't make this cutoff are mapped to a generic ``rare word" indicator. Lastly, we also map any numbers (that is numbers comprised of digits, not numbers written with words) to a generic ``number" indicator. Ultimately, this leaves us with a vocabulary size $|V|$ of 29,359. To estimate a bigram model, we simply need to compute the counts in our training corpus. Specifically, the probability of some word $w_i$ conditional on its preceding word $w_{i-1}$ is given by:
\begin{equation}
p(w_i|w_{i-1}) = \frac{\textnormal{count}(w_{i-1},w_i)}{\textnormal{count}}
\end{equation}
However, its reasonable to expect that there might be bigrams in the development or test corpora that aren't observed in the training corpus. This is particularly likely given that for our vocabulary size, there are nearly 900 million unique bigrams and our training data contains just over 3 million observations. If this is the case, then any article with a bigram unobserved in the training corpus would be assigned a predicted likelihood of 0. Needless to say, this is very bad. In order to avoid this issue, we apply a technique called add-$\alpha$ smoothing. Add-$\alpha$ smoothing adds $\alpha$ to each cell in the probability table. Hence, after smoothing, no bigram, given a fixed vocabulary $V$, will ever have 0 probability. This changes our our estimated word probabilities to:
\begin{equation}
p(w_i|w_{i-1}) = \frac{\textnormal{count}(w_{i-1},w_i) + \alpha}{\textnormal{count} + \alpha |V|}
\end{equation}
Essentially, this smoothing is acting like a probablistic Robin Hood. It takes probability mass from the ``rich" and distributes it to the ``poor". Additionally, this smoothing actually has a nice Bayesian interpretation. For the bigram language model case, we can say that words follow a multinomial distribution conditional on its preceding word. The smoothed language model is equivalent to the posterior of horizontally stacked multinomial likelihoods with a symmetric Dirichlet prior with a parameter value of $\alpha$.

@@

Here we see exactly what we should expect. The NALL on the training corpus is lower than the NALL on the validation corpus. Furthermore, the NALL on our training corpus monotonically increases as we increase $\alpha$ while the NALL on the validation corpus initially decreases, hits a minimum, and starts to increase. For our language model, the optimal $\alpha$ is 0.004. This level of smoothing produces the following NALLs for our 3 corpora:

@@

One criticism of standard n-gram language models is that they are rather sensitive to the training data. Suppose that two words generally have fairly interchangeable use cases (that is the sets of words that tend to precede these words tend to be highly similar). For example, consider ``coffee" and ``tea". Further, suppose that for whatever reason, the phrase ``drink coffee" is predominantly featured in the training data but ``drink tea" is not. Then a standard bigram language model would assign a very low probability to ``tea" if it were preceded by the word ``drink" even if the training data contains many instances in which they are fairly interchangeable (``buy coffee/tea", ``brew coffee/tea", etc.).

Hence, we build word vector n-gram neural network models to see if we can achieve better performance. By using dense word vectors rather than one-hot encodings to represent words, we're able to capture the underlying similarity of words and their meanings. For example, ``coffee" is more similar to ``tea" that it is to ``car". Rather astonishingly, these word vector values can be trained through back propagation just like the weights in a neural network as long as they're randomly initiated!

 In particular, we train both a bigram neural network language model and a trigram neural network language model. In both cases, we use word vectors of length 10 meaning that the input layer of the bigram NN is a 10 (a single word) and the input layer of the trigram is 20 (2 words concatenated together). This input layer then maps to the single hidden layer with 10 hidden units through a tanh activation function. Finally, the hidden layer is then connected to the output layer which has size $|V|$ through a softmax activation function. We use stochastic adaptive gradient descent so that the errors can be properly backpropagated to the word vectors. We purposefully keep our network relatively small since our data set is rather large and training even this small neural network takes considerable time.

@@

 We adapted our code so it saved a useable snapshot of the network after each full pass through the data. Unfortunately, as a result of an oversight in the code, we forgot to store the NALL of training data in this snapshot. All is not lost since we remembered to at least store the NALL on the validation corpus. We are highly confident our neural network is working as intended since we see in Figure \ref{fig:perp_hist} that the NALL decreases on every pass through the validation corpus.

@@

It's worth taking the time to discuss Figure \ref{fig:mse_plot}, which shows the training and holdout MSE for various values of $\lambda$, in slightly more depth. There are a few things in this plot worth discussing. First, note that the validation MSE is consistently higher than the training data MSE, which is consistently higher than the test data MSE. Given the (relatively) small size of our dataset (6,682 observations), this is likely due to the sampling we used to separate our data into training, validation, and test data. However, we don't expect this effect the validity of our cross validation.

@@

There is ample room to improve our sentiment analysis methodology. The training dataset generated using mechanical turk is relatively small (200 training articles). Because of this, there may exist many words that have a strong probability of appearing in an article conditional on sentiment that simply do not appear in our training set. With more time (and money!), we could label more data and improve the accuracy of our model. Furthermore, the model currently skips words that do not occur in the training corpus. An extension of our Naive Bayes implementation could use a method such as Laplace smoothing, so as to not simply ignore words we haven't seen in our training data.
